Automated Verification Applied to Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation

### 2019-07-22 Tex/Old text/chapters/chapter04.tex
thesis updated text

@@

%In order to evaluate the proposed verification method and its performance, we have considered five case studies, three verification engines in different configurations, and also compared the results to the HOMER Pro tool. All the experiments were performed with a timeout of 14,400 seconds. %Every dweller, who owns a PV system, was interviewed 
%
All experiments were conducted on an otherwise idle Intel Xeon CPU E5-4617 (8-cores) with 2.90 GHz and 64 GB of RAM, running Ubuntu 16.04 LTS 64-bits. The setup of HOMER Pro v3.13.1 was an Intel Core i5-4210 (4-cores), with 1.7 GHz and 4 GB of RAM, running Windows 10. The experiments were performed with timeout of 240 minutes.

Verification engine ESBMC, version v6.0.0 was used with the SMT solver Boolector version 3.0.1~\cite{Brummayer}\footnote{Command-line: \$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-unwind 100 -\phantom{}-boolector}; and an alternative ESBMC v6.0.0 was used with the SMT incremental mode\footnote{Command-line: \$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-unwind 100 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard -\phantom{}-z3} enabled; % with the goal of reducing memory usage; we have also used 
with SMT solver Z3 version 4.7.1~\cite{DeMoura}.

@@

% enabled with the goal of reducing memory usage; 
%The experiments were performed without a predefined timeout.
Verification engine CPAchecker 1.8 was used \footnote{Command-line: \$ scripts/cpa.sh -heap 64000m -stack 10240k -config config/bmc-incremental.properties -spec config/specification/sv-comp-reachability.spc filename.c}, with the SMT solver MathSAT version 5.5.3~\cite{mathsat5}. An alternative CPAchecker configuration was tried as well, using BMC k-induction option, but without improvements of performance or soundness in the results (so it is not reported here).
% enabled with the goal of reducing memory usage; 
 %The experiments were performed without a predefined timeout.

@@

However, running the same version of ESBMC but using incremental solving with Z3, the experimentation returned SAT or UNSAT to all the case studies.

Related to the cases that use a 700 W PV system (house 1, house 2, house 3, and house 4), ESBMC could not reach an error in all the four houses and the execution time took from 04 m 27 s to 05 m 08 s. Fig.~\ref{fig:esbmcverifhouse1} shows the output report of the tool, with the absence of fail at the end of analysis.

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{esbmcverifh1.png}
\centering
\caption{Report generated by ESBMC (with Z3 solver) after validation of House 1.}
\label{fig:esbmcverifhouse1}
\end{figure}

However the 1,200 W PV system (house 5) failed (SAT) in line 337 of the code, thereby indicating that the system is \textit{incorrectly} sized; in particular, the counterexample provided by the verification engine indicated that the nominal current from the charge controller is less than the minimum current demanded by the PV system, therefore the equipment chosen is not suitable to meet the design requirements. Fig.~\ref{fig:esbmcverifhouse5} shows the report issued by the tool, with red highlights for the result, showing the incompatibility of the sized charge controller and the minimum necessary. This verification took less than 1 s to be performed, as indicated in the last line of Table~\ref{cases}, and it is faster than the previous analysis because ESBMC stops during the sizing check, which is in line 3 of Algorithm~\ref{alg:verification-algorithm}, and does not perform the rest of verification code.

@@

Concerning the CBMC tool, similar results were obtained, but with some slower time than ESBMC. The experimentation returned SAT or UNSAT to all the case studies. Related to the 700 W PV systems (house 1, house 2, house 3, and house 4), the tool could not reach an error in all the four houses and the execution time took from 18 m 36 s to 19 m 02 s.

Fig. \ref{fig:cbmcverifhouse1} shows part of report produced by CBMC when validating the house 1. The file produced by CBMC is 160 lines long and 12.4 kB of size. The author highlighted the result, which shows that was not found design flaws within the bound $k$ of 100.

@@

However the 1,200 W PV system (house 5) failed (SAT) in line 337 of the code; with the same counterexample presented by ESBMC. This verification took less than 1 s to be performed as well. Fig.~\ref{fig:cbmcverifhouse5} shows the (edited) report produced, with the highlights about the flaw found: the minimum controller current needed to the PV systems is not compatible with the chosen controller of the sized system.

@@

Finally, the CPAchecker tool presented some different results. Even using two different configuration possibilities, as described in Section~\ref{sec:setup}, the verification engine presented an UNKNOWN result for all the 700 W systems. This is because the \textit{time out} limit was reached, i.e., after 4 hours of execution the tool was unable to decide if the verification was SAT or UNSAT. However, when verifying the 1,200 W PV system, the tool presented a SAT message equal to the other engines.

@@

Screens and comments about the simulation tool.
The same five case studies were evaluated by HOMER Pro (RQ2). The simulation results showed that the project restrictions were met by four 700 W PV systems (house 1, 2, 3 and 4), without any indication of sizing error or even performance related issues. The case study that was unsuccessful during simulation was the 1,200 W (house 5); however, without any indication about the failures of this PV system (RQ2). All the simulations took less than 5 seconds (each) to be performed by HOMER Pro. %Fig. \ref{fig:homerscreen} shows one of the screens presented by HOMER Pro software, specifically to house 1.
%
%\begin{figure}[h]
%\includegraphics[width=0.8\textwidth]{homer.png}
%\centering
%\caption{HOMER simulation screen.}
%\label{fig:homerscreen}
%\end{figure}
%

There were no divergence of results for the houses 1, 2, 3 and 4 w.r.t. our proposed approach, it is evident that the information collected from the dwellers and from the monitoring systems indicate that our approach provides the correct evaluation of the PV system, thus answering RQ2. House 5 presented flaws from all tools (automated verified or simulation); however, only automated verification approaches indicated which design error was responsible for the flaw (charge controller specification), further answering RQ2.
%
%Note that a PV design always uses daily average values of sun hours to each site, with impact in the PV components. Those hours are based on historical data and, in field, it is not unusual to find days where that number of hours was not reached due to weather conditions. The season has impact since the case studies are from the rain forest, where clouds are always present. As a result, the identified flaws in houses 1, 2, 3, and 4, are justified once again.
%
%We have evaluated five case studies in total using the HOMER Pro tool and our automated verification tool. Related to the HOMER Pro, the simulation, based on NASA data from the deployed systems (temperature and solar irradiance), shows that the restrictions were met by four 700 W PV systems (house 1, 2, 3 and 4), without any indication of sizing error or even performance. The case study that was unsuccessful during simulation was the 1,200 W (house 5); however, without any indication about the failures of this PV system. All the simulations took less than 5 seconds (each) to be performed.
%
%However, related to the results of the automated verification: (a) the 1,200 W PV system (house 5) failed during the sizing check. The number of panels was \textit{incorrect}; in particular, the counterexample provided by our verification method indicated 3 panels in parallel and the sized project has 2 in series and 2 in parallel. That verification took 63.3 hours to be performed. Surveying the owner of the 1,200 W system it was identified that, in fact, the system mostly of the time do not met the battery autonomy (mainly when all the loads are turned on). That behavior is expected because the system was purchased as an off-the-shelf solution and not as a specific design for the electrical charges of the house; (b) Related to the four 700 W PV systems, just one verification finished its analysis (house 1) considering the time-out of 432 h of computing. The sizing check was successful during automated verification, but there was found flaw related with the battery autonomy, when SOC reached levels below of 75\%. The automated verification identified the flaw right after the first night-discharge cycle, before the solar system start to recharge the batteries. The proposed tool took 409.3 hours to find this error at house 1. This possible flaw was confirmed with the dweller that uses the system: at least once or twice a mouth is usual the system to turn off, normally during raining days or with more clouds in the sky, and after the sun rises the system returns to normal operation. Related to houses 2, 3 and 4, it was considered that the automated verification had a time-out condition, with no conclusive results.

@@

In order to validate the possible flaw from house 5, we have surveyed the owner of the 1,200 W system. We identified that, in fact, the system does not meet the battery autonomy when all loads are turned on, and this was double checked with the monitoring system from the charge controller, which showed that the maximum power or surge power were not exceeded, thus affirming RQ1; this behavior is expected since the system was purchased as an off-the-shelf solution and not as a customized design for the electrical charges of the house. The same process of validation was done to the houses 1, 2, 3 and 4, which use the 700 W PV systems: from July of 2018 to March 2019, a monthly visiting was performed to apply surveys to the dwellers and to collect data from a local monitoring system: not every month were reported some energy interruption of the PV systems. However, even when one interruption is reported in a month, this represents around 3.33\% of interruption for the entire period ($1/30$), which indicates 96.97\% of availability of the PV system %($96.67\% = 100\%-3.33\%$) 
and it is in accordance to what was described in Section~\ref{sec:availability}, because the type of electrical load of the houses is not critical; this situation is considered an energy interruption, but is not considered a system flaw, further affirming RQ1.
%\begin{figure}[h]
%\includegraphics[width=0.65\textwidth]{loadcurve.png}
%\centering
%\caption{Five weeks monitored load curve from House 1.}
%\label{fig:loadcurve}
%\end{figure}

%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\section{Threats to Validity}
%---------------------------------------------------
We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified three threats to the validity of our results that can further be assessed.

\textit{Model precision:} each component of the PV system is mathematically modeled. %, and the precision of the proposed method depends on the precision of that particular model. 
The adoption of more complex models, or even an evaluation in a PV laboratory to validate the model could add more reliability to the results.

@@

>>>>

Worth to mention that the load curve represents a array of 24 integer numbers for every hour of the day (instant power in Watts) and it was estimated after visiting and survey applied in June of 2017, before the houses were electrified.

@@

%In order to evaluate the proposed verification method and its performance, we have considered five case studies, three verification engines in different configurations, and also compared the results to the HOMER Pro tool. All the experiments were performed with a time out of 14,400 seconds. %Every dweller, who owns a PV system, was interviewed 
%
All experiments were conducted on an otherwise idle Intel Xeon CPU E5-4617 (8-cores) with 2.90 GHz and 64 GB of RAM, running Ubuntu 16.04 LTS 64-bits. The setup of HOMER Pro v3.13.1 was an Intel Core i5-4210 (4-cores), with 1.7 GHz and 4 GB of RAM, running Windows 10. The experiments were performed with time out of 240 minutes.

Verification engine ESBMC, version v6.0.0 was used with the SMT solver Boolector version 3.0.1~\cite{Brummayer}\footnote{Command-line: \$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-unwind 100 -\phantom{}-boolector}; and an alternative ESBMC v6.0.0 was used with the SMT incremental mode\footnote{Command-line: \$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-unwind 100 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard -\phantom{}-z3} enabled; with SMT solver Z3 version 4.7.1~\cite{DeMoura}. % with the goal of reducing memory usage.

@@

Verification engine CPAchecker 1.8 was used \footnote{Command-line: \$ scripts/cpa.sh -heap 64000m -stack 10240k -config config/bmc-incremental.properties -spec config/specification/sv-comp-reachability.spc filename.c}, with the SMT solver MathSAT version 5.5.3~\cite{mathsat5}. An alternative CPAchecker configuration was tried as well, using BMC k-induction option, but without improvements of performance or soundness in the results (so it is not reported here).

@@

However, running the same version of ESBMC but using incremental solving with Z3, the experimentation returned was conclusive (SAT or UNSAT) to all the case studies.

Related to the cases that use a 700 W PV system (house 1, house 2, house 3, and house 4), ESBMC could not reach an error in all the four houses and the execution time took from 04 m 27 s to 05 m 08 s. Fig.~\ref{fig:esbmcverifhouse1} shows the output report of the tool for the house 1, with the absence of fail at the end of analysis. The file produced by ESBMC is 160 lines long and 12.4 kB of size. The author highlighted the result, which shows that was not found design flaws (SAT result or SUCCESS) within the bound $k$ of 100.

\begin{figure}[h]
\includegraphics[width=0.65\textwidth]{esbmcverifh1.png}
\centering
\caption{Report generated by ESBMC (with Z3 solver) after validation of House 1.}
\label{fig:esbmcverifhouse1}
\end{figure}

However the 1,200 W PV system (house 5) failed (SAT) in line 337 of the code, thereby indicating that the system is \textit{incorrectly} sized; in particular, the counterexample provided by the verification engine indicated that the nominal current from the charge controller is less than the minimum current demanded by the PV system, therefore the equipment chosen is not suitable to meet the design requirements.

Fig.~\ref{fig:esbmcverifhouse5} shows the report issued by the tool, with red highlights for the result, showing the incompatibility of the sized charge controller and the minimum necessary. The report generated by ESBMC was a text file of 14.1 kB of size and 353 lines of length. This verification took less than 1 s to be performed, as indicated in the last line of Table~\ref{cases}, and it is faster than the previous analysis because ESBMC stops during the sizing check, which is in line 3 of Algorithm~\ref{alg:verification-algorithm}, and does not perform the rest of verification code.

@@

Concerning the CBMC tool, similar results were obtained, but with some slower time than ESBMC. The experimentation returned SAT or UNSAT to all the case studies, i.e., it was conclusive to all houses.

Related to the 700 W PV systems (house 1, house 2, house 3, and house 4), the tool could not reach an error in all the four houses and the execution time took from 18 m 36 s to 19 m 02 s.

Fig. \ref{fig:cbmcverifhouse1} shows part of report produced by CBMC when validating the house 1. The file produced by CBMC is 270,512 lines long and 39 MB of size. The author highlighted the result, which shows that was not found design flaws within the bound $k$ of 100.

@@

However the 1,200 W PV system (house 5) failed (SAT) in line 337 of the code; with the same counterexample presented by ESBMC. This verification took less than 1 s to be performed as well. Fig.~\ref{fig:cbmcverifhouse5} shows the (edited) report produced, with the highlights about the flaw found: the minimum controller current needed to the PV systems is not compatible with the chosen controller of the sized system. The report file generated contained 217 lines and was 13 kB long.

@@

Finally, the CPAchecker tool presented some different results. Even using two different configuration possibilities, as described in Section~\ref{sec:setup}, the verification engine presented an UNKNOWN result for all the 700 W systems. This is because the \textit{time out} limit was reached, i.e., after 4 hours of execution the tool was unable to decide if the verification was SAT or UNSAT.

Figure here CPA house 1

However, when verifying the 1,200 W PV system, the tool presented a SAT message equal to the other engines.

Figure here CPA house 5

@@

Screens and comments about the simulation tool.
1) If you want to simulate a particular capacity for PV and battery, you can use the search space instead of HOMER Optimizer for sizing. You can either aggregate all the 3 PV panels together as a single PV component or add 3 PV components to represent the series configuration. For the batteries, you will have to find the equivalent capacity of the series and parallel configuration since HOMER considers only one battery component per simulation even if you add multiple battery components in the schematic.
2) You cannot set the autonomy of the batteries as HOMER controller will decide when it is economical to discharge the batteries during the simulation year.

The same five case studies were evaluated by HOMER Pro (RQ2). The simulation results showed that the project restrictions were met by four 700 W PV systems (house 1, 2, 3 and 4), without any indication of sizing error or even performance related issues. All the simulations took less than 5 seconds (each) to be performed by HOMER Pro. Fig. \ref{fig:homerscreen1} shows one of the screens presented by HOMER Pro software, specifically to house 1.

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{homer.png}
\centering
\caption{HOMER simulation screen.}
\label{fig:homerscreen1}
\end{figure}

The case study that was not possible to simulate was the 1,200 W (house 5). The reason is based on the fact that HOMER Pro do not allow to set up a battery autonomy; therefore, it was was not possible to obtain any indication about the failures of this specifically PV system with simulation (RQ2).

%
%Note that a PV design always uses daily average values of sun hours to each site, with impact in the PV components. Those hours are based on historical data and, in field, it is not unusual to find days where that number of hours was not reached due to weather conditions. The season has impact since the case studies are from the rain forest, where clouds are always present. As a result, the identified flaws in houses 1, 2, 3, and 4, are justified once again.
%
%We have evaluated five case studies in total using the HOMER Pro tool and our automated verification tool. Related to the HOMER Pro, the simulation, based on NASA data from the deployed systems (temperature and solar irradiance), shows that the restrictions were met by four 700 W PV systems (house 1, 2, 3 and 4), without any indication of sizing error or even performance. The case study that was unsuccessful during simulation was the 1,200 W (house 5); however, without any indication about the failures of this PV system. All the simulations took less than 5 seconds (each) to be performed.
%
%However, related to the results of the automated verification: (a) the 1,200 W PV system (house 5) failed during the sizing check. The number of panels was \textit{incorrect}; in particular, the counterexample provided by our verification method indicated 3 panels in parallel and the sized project has 2 in series and 2 in parallel. That verification took 63.3 hours to be performed. Surveying the owner of the 1,200 W system it was identified that, in fact, the system mostly of the time do not met the battery autonomy (mainly when all the loads are turned on). That behavior is expected because the system was purchased as an off-the-shelf solution and not as a specific design for the electrical charges of the house; (b) Related to the four 700 W PV systems, just one verification finished its analysis (house 1) considering the time-out of 432 h of computing. The sizing check was successful during automated verification, but there was found flaw related with the battery autonomy, when SOC reached levels below of 75\%. The automated verification identified the flaw right after the first night-discharge cycle, before the solar system start to recharge the batteries. The proposed tool took 409.3 hours to find this error at house 1. This possible flaw was confirmed with the dweller that uses the system: at least once or twice a mouth is usual the system to turn off, normally during raining days or with more clouds in the sky, and after the sun rises the system returns to normal operation. Related to houses 2, 3 and 4, it was considered that the automated verification had a time-out condition, with no conclusive results.

@@

There were no divergence of results for the houses 1, 2, 3 and 4 w.r.t. our proposed approach, it is evident that the information collected from the dwellers and from the monitoring systems indicate that our approach provides the correct evaluation of the PV system, thus answering RQ2. House 5 presented flaws from all tools (automated verified or simulation); however, only automated verification approaches indicated which design error was responsible for the flaw (charge controller specification), further answering RQ2.

In order to validate the possible flaw from house 5, we have surveyed the owner of the 1,200 W system. We identified that, in fact, the system does not meet the battery autonomy when all loads are turned on, and this was double checked with the monitoring system from the charge controller, which showed that the maximum power or surge power were not exceeded, thus affirming RQ1; this behavior is expected since the system was purchased as an off-the-shelf solution and not as a customized design for the electrical charges of the house. The same process of validation was done to the houses 1, 2, 3 and 4, which use the 700 W PV systems: from July of 2018 to March 2019, a monthly visiting was performed to apply surveys to the dwellers and to collect data from a local monitoring system: not every month were reported some energy interruption of the PV systems. However, even when one interruption is reported in a month, this represents around 3.33\% of interruption for the entire period ($1/30$), which indicates 96.97\% of availability of the PV system ($96.67\% = 100\%-3.33\%$) 
and it is in accordance to what was described in Section~\ref{sec:availability}, because the type of electrical load of the houses is not critical; this situation is considered an energy interruption, but is not considered a system flaw, further affirming RQ1.

%\begin{figure}[h]
%\includegraphics[width=0.65\textwidth]{loadcurve.png}
%\centering
%\caption{Five weeks monitored load curve from House 1.}
%\label{fig:loadcurve}
%\end{figure}
%
%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\section{Threats to Validity}
%---------------------------------------------------

We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified three threats to the validity of our results that can further be assessed.

\textit{Model precision:} each component of the PV system is mathematically modeled, and the precision of the proposed method depends on the precision of that particular model. 
The adoption of more complex models, or even an evaluation in a PV laboratory to validate the model could add more reliability to the results.

@@

At this Chapter it was performed the comparative among verifiers with the proposed automated verification using model checking, which showed that 'incremental' ESBMC had the best overall performance. CBMC presented the same results, however with a worse time to result, and CPAchecker was not conclusive (house 1, house 2, house 3, and house 4) because the \textit{time out} was not enough to obtain some result.

A comparative with a specialized simulation tool was performed as well, however some limitations of the tool, mainly related to not allow battery autonomy setup, restricted the comparative between automated verification tools versus simulation tools.

Based on the fact that all case studies were deployed at the field, with regular visiting to perform interview with the dwellers, and with monitoring system in some of the houses (house 1, house 2, house 3, and house 4), it was possible to compare the computer obtained results with the real world employment of the stand-alone solar PV systems. The final conclusion was that the proposed tool is sound and with an acceptable performance.
